\bfseries\LARGE\singlespacing{Transmission of respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-10-03 doc/paper/manuscript.tex

\noindent We computed the daily number of detected coughs and compared it between study conditions. We estimated the reduction in the number of coughs using a Bayesian Negative Binomial regression model (Supplementary Text~\zref{sec:aud-regression-model}) with time in class as the model offset, adjusting for class and weekday effects, the number of students in class, the air change rate, and the cumulative number of respiratory cases. In addition, we used a Bayesian Negative Binomial hierarchical (random effects) regression model to estimate the association between the number of coughs and the number of positive saliva test results for respiratory viruses. \medskip

@@

Technical reasons for the differences in airborne detection are unlikely. The bioaerosol sampling devices and laboratory procedures were the same in both studies and no technical issues were observed. Instead, airborne detection could have been influenced by environmental factors. Although temperature (between 19 and 25$^{\circ}$C) and relative humidity (between 25 and 50\%) were similar in both studies, the same levels can have a different effect on airborne survival of respiratory viruses. For example, for rhinovirus and adenovirus, survival increases with relative humidity, while the reverse relationship has been documented for influenza and SARS-CoV-2\cite{Tellier2009JTRSI,Ahlawat2020AAQR,Biryukov2020mS,Karim1985CJM,Davis1971AM}. A change in ventilation conditions could have had a profound effect, but CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Although lower ventilation was partially offset by a greater reduction in aerosol and particle concentrations from air cleaners, the removal of particles through air cleaners may be slower than through ventilation.

@@

Our study indicated a small, potential benefit of air cleaners, but other infection control measures such as universal mask wearing appear to be more effective\cite{Banholzer2023PLoSMed,Heinsohn2022,Gettings2021,Dharmadhikari2012AJRCCM,Leung2020NatMed,Milton2013PLoSPathogens}. There may be several reasons for this. Compared to \eg mask wearing, air cleaners cannot influence transmission outside the classrooms. Furthermore, previous work suggests that prolonged close contact may be needed for transmission\cite{Leung2020NatMed,Brankston2007LancetID}. However, compared to face masks, air cleaners are unable to prevent such close range high particle density aerosol transmission. Evidence from our previous\cite{Banholzer2023PLoSMed} and this study further suggest that air cleaners are more effective in the removal of larger particles ($>5\mu$m). Yet, a larger proportion of respiratory viruses are carried in the smaller particles ($\leq5\mu$m)\cite{Fennelly2020}. Finally, classroom activity, airflow and other unobserved, confounding factors make it difficult to assess the effects of air cleaners. Nevertheless, it is important to assess their effectiveness in a real-life setting, and compare with the hypothetical impact suggested by simulation studies\cite{Lindsley2021,Cortellessa2023Build}.

>>>>

\noindent We computed the daily number of detected coughs and compared it between study conditions. We estimated the reduction in the number of coughs using a Bayesian Negative Binomial regression model (Supplementary Text~\zref{sec:aud-regression-model}) with time in class as the model offset, adjusting for class and weekday effects, the number of students in class, the air change rate, and the cumulative number of respiratory cases. In addition, we used a Bayesian Negative Binomial hierarchical (random effects) regression model to estimate the association between the number of coughs and the number of positive saliva test results for respiratory viruses. For this analysis, we only used the days were saliva testing was performed and coughs could be detected because people were in the classroom (27~days in total).  \medskip

@@

Technical reasons for the differences in airborne detection are unlikely. The bioaerosol sampling devices and laboratory procedures were the same in both studies and no technical issues were observed. Instead, airborne detection could have been influenced by environmental factors. Although temperature (overall between 19 and 25$^{\circ}$C) and relative humidity (overall between 25 and 50\%) were similar in both studies, the same levels can have a different effect on airborne survival of respiratory viruses. For example, for rhinovirus and adenovirus, survival increases with relative humidity, while the reverse relationship has been documented for influenza and SARS-CoV-2\cite{Tellier2009JTRSI,Ahlawat2020AAQR,Biryukov2020mS,Karim1985CJM,Davis1971AM}. A change in ventilation conditions could have had a profound effect, but CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Although lower ventilation was partially offset by a greater reduction in aerosol and particle concentrations from air cleaners, the removal of particles through air cleaners may be slower than through ventilation.

@@

Our study indicated a small, potential benefit of air cleaners, but other infection control measures such as universal mask wearing appear to be more effective\cite{Banholzer2023PLoSMed,Heinsohn2022,Gettings2021,Dharmadhikari2012AJRCCM,Leung2020NatMed,Milton2013PLoSPathogens}. There may be several reasons for this. Compared to \eg mask wearing, air cleaners cannot influence transmission outside the classrooms. Furthermore, previous work suggests that prolonged close contact may be needed for transmission\cite{Leung2020NatMed,Brankston2007LancetID}, or make transmission more likely despite prior vaccination or infection\cite{Lind2023NatCommun}. However, compared to face masks, air cleaners are unable to prevent such close range high particle density aerosol transmission. Evidence from our previous\cite{Banholzer2023PLoSMed} and this study further suggest that air cleaners are more effective in the removal of larger particles ($>5\mu$m). Yet, a larger proportion of respiratory viruses are carried in the smaller particles ($\leq5\mu$m)\cite{Fennelly2020}. Finally, classroom activity, airflow and other unobserved, confounding factors make it difficult to assess the effects of air cleaners. Nevertheless, it is important to assess their effectiveness in a real-life setting, and compare with the hypothetical impact suggested by simulation studies\cite{Lindsley2021,Cortellessa2023Build}.
